---
title: "AE 07: Multiple Linear Regression"
subtitle: "Credit Cards"
editor: visual
format:
  html:
    embed-resources: true
---

::: callout-important
-   Open [RStudio](https://rstudio.collegeofidaho.edu/) and create a subfolder in your AE folder called "AE-07".

-   Go to the [Canvas](https://cofi.instructure.com/courses/17093/assignments/202838) and locate your `AE-07` assignment to get started.

-   Upload the `ae-07.qmd` file into the folder you just created.
:::

## Packages + data

```{r load-packages}
#| message: false
library(readr)
library(tidyverse)
library(ggformula)
library(broom)
library(mosaic)
library(knitr)
library(ISLR2)
library(GGally)
library(yardstick)
```

The data for this AE is from the `Credit` data set in the **ISLR2** R package. It is a simulated data set of 400 credit card customers. We will focus on the following variables:

**Response**

-   `Limit`: Credit limit

**Predictors**

-   Everyone:

    -   `Rating`: Credit Rating

-   Group 1:

    -   `Age`
    -   `Own`

-   Group 2:

    -   `Balance`
    -   `Student`

-   Group 3:

    -   `Education`
    -   `Region`

-   Group 4:

    -   `Income`
    -   `Married`

## Analysis goal

The goals of this analysis is to fit a multiple linear regression model that that predicts a person's credit limit.

## Exercise 0

Look at the documentation for the data using `?Credit` and Google to figure out what each of your variables represents. What is a credit rating and what is a credit limit as it applies to a credit card? The primary credit rating in the US is called a FICO score. Use `favstats` on `Rating`. Do you think that `Rating` corresponds to the borrower's FICO score?

```{r}
favstats(~Rating, data = Credit)
```

Without writing any more code (i.e. just using what you know about the world):

-   Do you think `Rating` will be a good predictor of `Limit`?

    Yes, it is.

-   In addition to `Limit`, everyone has been assigned two additional variables, one quantitative and one categorical. Do you think these two variables will be predictive of `Limit`?

    Yes both of our variable which in this case are Education and Region are predictive of Limit.

## Exercise 1

Use the function `ggpairs` from the `GGally` package (already loaded) to create a matrix of plots and correlations for your four variables of interest. Note that you will have to use `select` to select the four columns you are interested in. Which variable do you think will be the best predictor of `Limit`?

```{r}
Credit |> 
  select(Limit, Rating, Education, Region) |>
  ggpairs()
```

## Exercise 2

Fit a linear model with just `Rating` as the predictor and get the p-value associated with it's coefficient. Is it statistically significant?

```{r}
credit_fit <- lm(Limit ~ Rating, data = Credit)
tidy(credit_fit) |>
  kable(digits = 5)
```

Yes! it means there is high evidence that there is a relationship between Limit and Rating.

## Exercise 3

Fit a linear model with just your second quantitative predictor and get the p-value associated with it's coefficient. Is it statistically significant?

```{r}
credit_fit <- lm(Limit ~ Education, data = Credit) 
tidy(credit_fit) |> kable(digits = 5)
```

No. It is because our p values is greater than 0.05 which shows high evidence to null hypothesis.

## Exercise 4

Fit a model with both `Income` and your second quantitative variable as predictors. Find a spot on the white board to write down an equation representing the fitted model. How do the coefficients and p-values of `Income` and `Rating` compare to those in the two models above? Discuss what you see and the possible reasons you see them.

```{r}
credit_fit <- lm(Limit ~ Education + Rating, data = Credit) 
tidy(credit_fit) |> kable(digits = 5)
```

P value for education became much more significant. Also the slope of rating is high highlighting how important of a factor is credit card rating in measuring credit card limit.

## Exercise 5

Interpret all coefficients in the model.

-   Someone with a 0 years of education and 0 rating in credit cars, will have -608 credit score which is not possible because the lowest credit score we have ins 93.

-   For a unit increase in years of education, the limit increases by 4.79 keeping ratings constant.

-   For a unit increase in ranking of credit card, the limit increases by 14.87 keeping education constant.

## Exercise 6

Use `head` to look at the first observation in the data set. Use addition, subtraction, multiplication, and division to figure out what credit limit your model would predict for a person with the same characteristics of this person.

```{r}
head(Credit, n=1) |> kable()


credit_limit <- -608.52 + 4.79*11  + 14.87*283

# 3652.38
```

## Exercise 7

Now use the `predict` function to verify the value you computed in the previous exercise and compute a 90% prediction interval. How would you interpret this interval in context?

```{r}
predict_data <- tibble(Rating = 283, Education = 11)
predict(credit_fit, predict_data, interval = "prediction", level = 0.90) 
```

90% of the times, the credit_fit falls between lower and upper level as shown.

```{r}
new_borrow <- head(Credit, n=10)
predict(credit_fit, new_borrow, interval = "prediction", level = 0.9)
```

## Exercise 8

Fit a linear model predicting `Limit` from **ONLY** your categorical variable. What is the reference level?

The reference level is 4881.5960.

```{r}
region_fit <- lm(Limit~ Region, data = Credit)
tidy(region_fit) |> kable()
```

## Exercise 9

WITHOUT WRITING ANY CODE except for addition, subtraction, multiplication, and addition, what would the model predict the average `Limit` to be for each of the level of your categorical variable?

```{r}
4881.5960

4881.5960 - 153.1336 #south

4881.5960 - 273.7724 #west
```

## Exercise 10

Compute the average `Limit` for each level in your categorical variable. Hint: you can use "formulas" (i.e. `repsonse ~ explanatory`) in the `mean` function from the `mosaic` package. How do these answers compare to your answers from Exercise 9?

```{r}
mean(Limit ~ Region, data = Credit)
```

```{r}
library(coursekata)
gf_jitter(Limit ~ Region, data = Credit) |>
  gf_model(region_fit)
```

## Exercise 11

Fit a linear model using your quantitative and your categorical variable as predictors (don't use `Rating`). Use the function `plotModel` to make a nifty plot for your model. Your syntax should look like `plotModel(model_name)`.

```{r}
original_fit <- lm(Limit ~ Education + Region, data = Credit) 
tidy(original_fit) |> kable(digits = 5)
plotModel(original_fit)
```

## Exercise 12

Your model contains one line for every level of your categorical variable. On the board, write out the equations for two of them.

Done

## Exercise 13

How do you think the plot above will change if you add in an interaction term between your variables? **AFTER** thinking about it, add in an in the interaction term and plot the model using `plotModel`.

```{r}
interaction_fit <- lm(Limit ~ Education + Region + (Education * Region), data = Credit) 
tidy(interaction_fit) |> kable(digits = 5)
plotModel(interaction_fit)
```

## Exercise 14

Your model contains one line for every level of your categorical variable. On the board, write out the equations for two of them.

LIMIT for East = (7159.8) - 168.4 \* Education

LIMIT for South= (7159.8-3078.4) +(-168.4435+216.98) \* Education

LIMIT for West = (7159.8 - 2499.4) + (-168.4435+164.58) \* Education

## Exercise 15

Fit a model with your quantitative predictor, `Rating`, and an interaction term. Write your model on the board.

```{r}
interaction_fitr <- lm(Limit ~ Rating + Education + (Rating * Education), data = Credit) 
tidy(interaction_fitr) |> kable(digits = 5)
plotModel(interaction_fitr)
```

## Exercise 16 (Optional)

Look back at all of the models that you've fit. If you compare the p-values for the same variable across multiple models, what do you notice?

## Exercise 17 (Optional)

Note that this data set only considers borrowers who have actually been granted loans. How does this impact the generalizability of our analysis?

## To submit the AE

::: callout-important
-   Render the document to produce the HTML file with all of your work from today's class.
-   Upload your QMD and HTML files to the Canvas assignment.
:::
